### 2021-07-22 paper/paper.tex
WIP: assessment of Julia

In traditional monolithic code bases, it is often necessary to implement a new
feature for all combinations of possible (compile time) options. This effort
can make it hard to experiment with new ideas. In contrast, Julia's dynamic
nature and just-ahead-of-time compiler allow us to implement only the features
that are strictly necessary. For example, if a user wants to simulate a new
physics model only on Cartesian grids with smooth solutions, there is no need
to implement anything for curvilinear coordinates or shock capturing approaches.

@@

@@

\todo{TODO %TODO
\begin{itemize}
  \item Packages can really be used together at low (or no) cost
  \item Reproducibility
  \item Good software development practices (CI, tests, version management)
  \item Relatively easy to integrate existing C libraries (p4est, HDF5)
  \item more nice features
  \item Can be extended: Multi-physics with 350 lines of code
  \item Compilation time can be annoying but also allows cool features and performance \cite{schlottkelakemper2021purely}
  \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
\end{itemize}}

@@

\todo{ % TODO
\begin{itemize}
  \item Prepare repro repo including scripts for \trixi and Fluxo
  \item Cite \cite{schlottkelakemper2021purely,revels2016forward,ranocha2021sbp}
  \item The title of our talk at JuliaCon is
  ``Adaptive and extendable numerical simulations with Trixi.jl''.
  Shall we use something like
  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
  A case study of Julia for scientific computing'' for this paper?
  \item Restructure non-conservative stuff and add MHD benchmarks?
  \item Shall we re-create the schematic overview in tikz?
  \item Shall we compare \trixi to other open source frameworks such as MFEM, NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...?
\end{itemize}}

\clearpage
\section*{Some guidelines}
\label{sec:some_guide}

Full paper: 5--10 pages.
\begin{quote}
  Compared to an extended abstract, a full paper presents more of the background
  and context motivating the work. It compares the work to other approaches taken
  in the field and gives some additional insights on the conference contribution.
  Use cases back up the work by showing how it can be used.
\end{quote}

The following notes may help you achieve the best effects with the
\verb juliacon  class file.

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Enunciations}
\label{subsub:enunciation}

The \verb juliacon   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use:
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

% \vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

>>>>

In traditional monolithic code bases, it is often necessary to implement a new
feature for all combinations of possible (compile time) options. This effort
can make it hard to experiment with new ideas. In contrast, Julia's dynamic
nature and just-ahead-of-time compiler allow us to implement only the features
that are strictly necessary. For example, if a user wants to simulate a new
physics model only on Cartesian grids with smooth solutions, there is no need
to implement anything for curvilinear coordinates or shock capturing approaches.
In addition, it allows to extend \trixi step by step with new capabilities.
For example, there are ongoing efforts to incorporate summation by parts finite
difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi. This becomes feasible since not every
feature combination needs to be considered and it is totally fine to start with
simple variants on a subset of the available meshes.

@@

The code and detailed information necessary to reproduce these numerical
experiments is available in the accompanying repository \cite{ranocha2021adaptiveRepro}.

@@

Julia was designed from scratch to be useful for numerical computing
\cite{bezanson2017julia}. Its usefulness has been demonstrated for example for
optimization \cite{dunning2017jump}, data science including
big data\footnote{Celeste project: \url{https://github.com/jeff-regier/Celeste.jl},
see \url{https://juliacomputing.com/case-studies/celeste/} (accessed 2021-07-22)},
machine learning \cite{innes2018fashionable}, and scientific machine learning
(SciML) \cite{pal2021opening}. There are also mature libraries for ODE problems
\cite{rackauckas2017differentialequations}, small scale spectral approximations
\cite{olver2014practical}, and classical finite element methods \cite{badia2020gridap}.
In addition, there are some packages dedicated to solve specific time-dependent
PDEs such as \cite{ramadhan2020oceananigans, constantinou2021geophysicalflows}.
However, there does not seem to be a general framework of high-order methods
or hyperbolic PDEs in Julia. Thus, \trixi is an ideal candidate for a case study
of Julia for simulation-focused scientific computing, an area where codes written
in classical programming languages such as C and Fortran are still dominant
\cite{krais2021flexi, parsani2021ssdc}.

\subsection{What works well}

\subsubsection{Julia encourages good software development practices}

Many scientists implementing numerical methods often have no real experience or
education in software development. Julia and its ecosystem support these
researches by making it easy to set up unit and regression tests, since a
testing framework is included in the standard library and default package
generation setups prepare the file structure accordingly. This allows test
driven development and continuous integration (CI), which makes code restructuring
and optimization feasible.

In addition, the package manager encourages a unified form of semantic versioning
across the Julia ecosystem, which lays the basis for the following three
observations.

\subsubsection{It is easy to set up reproducible numerical experiments}

The package manager makes it easy to reproduce the exact environment, including
binary dependencies. We have used this feature for all of our papers based on
\trixi \cite{schlottkelakemper2021purely, ranocha2021preventing}, including this
one \cite{ranocha2021adaptiveRepro}.
This facilitates code sharing and reproducible research in computational science,
which is arguably important but not mainstream yet \cite{barnes2010publish,
donoho2010invitation, leveque2013top}.

\subsubsection{External C libraries can be integrated relatively easily}

\cite{byrne2021mpi}

\todo{TODO: p4est \cite{burstedde2011p4est} (HDF5)} %TODO

\subsubsection{Packages can really be used together at low (or no) cost}

\subsubsection{The code base is simple enough to be useful for students}

\subsubsection{Existing features can be extended and combined easily}

Multi-physics with 350 lines of code \cite{schlottkelakemper2021purely}
\todo{TODO} %TODO

\subsubsection{\trixi can be extended from outside}

\subsection{What is still difficult or unknown}

\todo{TODO %TODO
\begin{itemize}
  \item Compilation time can be annoying (serial compiler!) but also allows cool features and performance
  \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
\end{itemize}}

@@

\todo{ % TODO
\begin{itemize}
  \item Full paper: 5--10 pages
  \item Prepare repro repo including scripts for \trixi and Fluxo
  \item Cite \cite{schlottkelakemper2021purely,revels2016forward,ranocha2021sbp}
  \item The title of our talk at JuliaCon is
  ``Adaptive and extendable numerical simulations with Trixi.jl''.
  Shall we use something like
  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
  A case study of Julia for scientific computing'' for this paper?
  \item Restructure non-conservative stuff and add MHD benchmarks?
  \item Shall we re-create the schematic overview in tikz?
  \item Shall we compare \trixi to other open source frameworks such as MFEM, NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...?
\end{itemize}
\begin{quote}
  Compared to an extended abstract, a full paper presents more of the background
  and context motivating the work. It compares the work to other approaches taken
  in the field and gives some additional insights on the conference contribution.
  Use cases back up the work by showing how it can be used.
\end{quote}}
